\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
\justifying


The recent emergence of smartphones has led to a XX in the amount of mobile applications available for all existing platforms. In the app markets a wide variety of applications can be found, ranging from games or social networks to some more career-oriented. The result is a rapidly growing market devoted to their development with programming languages such as Java, Swift or Objective C.

However, very different platforms exist, owned by different companies. This means that once an app has been developed for one of these platforms, in order to run it on another one, the application has to be completely doveloped almost from scratch again.

In order to ease the task of porting an app to a different computing environment emerged the concept of hybrid mobile applications. Using web technologies such as HTML5, CSS and JavaScript, and an API, cross-platform mobile applications can be developed in a a quick and efficient manner.

Currently, there are two technologies for the development of hybrid mobile applications: Phonegap, used by a broad amount of developers, and the more recent React Native, which is still under development. In this work, the biggest and most important differences between both technologies will be explained, and in order to hightlight them, the same application will be developed for them both.

A React Native test scenario will be designed and implemented as well, to act as an introduction to this technology. Every step of the process will be thoroughly documented, from  installing the required software to running the app on a device.

\vspace{5mm}
\large{\textbf{Keywords: } Aplicación Móvil, React, Phonegap, JavaScript.}
\afterpage{\blankpage}